---
title: "Data"
execute:
  echo: false
  warning: false
  message: false
---

Below you will find the datasets used in this course. You can download them directly or use the provided links to load them into R/RStudio.

## False beginners

This dataset contains ... 

* **Download:** FalseBeginners.txt
* **Format:** tab delimited text file

#### Codebook

| Variable | Description |
|----------|-------------|
| `student` | ... |
| `School` | ... |
| `Class` | ... | 
| `Class` | ... |

#### Data Preview

```{r}
fb <- read.delim("FalseBeginners.txt") 
head(fb)
```

## Vocatives

dataset_ImPoliteVocatives_20230530.csv

This dataset contains ... 

* **Download:** dataset_ImPoliteVocatives_20230530.csv

#### Codebook

| Variable | Description |
|----------|-------------|
| `student` | ... |
| `School` | ... |
| `Class` | ... | 
| `Class` | ... |

#### Data Preview

```{r}
voc <- read.csv2("dataset_ImPoliteVocatives_20230530.csv") 
options(width = 60) # to print columns side by side
head(voc)
```

## Listening to accents

This dataset contains ... 

* **Download:** Listening_to_Accents_Comprehensibility_Accentedness.csv

#### Codebook

| Variable | Description |
|----------|-------------|
| `student` | ... |
| `School` | ... |
| `Class` | ... | 
| `Class` | ... |

#### Data Preview

```{r}
accent <- read.csv("Listening_to_Accents_Comprehensibility_Accentedness.csv") 
head(accent)
```

